---
title: "Prettymapr: Tools for rapid, nice-looking maps in R"
author: "Dewey Dunnington"
date: "April 1, 2017"
output: github_document
---

[![](http://cranlogs.r-pkg.org/badges/prettymapr)](https://cran.r-project.org/package=prettymapr)

Prettymapr automates the process of creating a scale bar and north arrow in any package that uses base graphics to plot in R, or provides parameters that help to draw scale bars and north arrows in other plotting environments. Bounding box tools help find and manipulate extents, and geocoding tools help plot locations on maps. Finally, there is a function to automate the process of setting margins, plotting the map, scale bar, and north arrow, and resetting graphic parameters upon completion.

## Installation

The prettymapr package is available on CRAN, and can be installed using `install.packages("prettymapr")`. The manual is also [available on the CRAN website](https://cran.r-project.org/web/packages/prettymapr/prettymapr.pdf)

## Bounding box tools

The concept of a bounding box is integral to the [rosm](https://cran.r-project.org/package=rosm) and [rcanvec](https://cran.r-project.org/package=rcanvec) packages (as well as some others), so there's a few tools to create/manipulate them here. I can never remember the proper way to create a bounding box using `sp::bbox()`, so the `makebbox(n, e, s, w)` function is included to make this process a little easier (I can somehow always remember 'never eat shredded wheat' for the cardinal directions). The easiest way to get these numbers is from the [Open Street Maps Export page](http://www.openstreetmap.org/export).

```{r}
library(prettymapr)
makebbox(50.1232, -122.9574, 50.1035, -123.0042)
```

It is, of course, easier to search for bounding boxes by human-readable location (google maps style), although this requires some confidence in the internet to recognize your location (the default `source` is the [PickPoint API](https://pickpoint.io/), although you can pass `source="google"` to use the [Google Maps API](https://developers.google.com/maps/), or `source="dsk"` to use the [Data Science Toolkit](http://www.datasciencetoolkit.org/)).

```{r, message=FALSE}
searchbbox("alta lake, british columbia")
```

If you pass a vector into `searchbbox()` it will return the bounding box that contains all of them.

```{r, message=FALSE}
searchbbox(c("halifax, ns", "moncton, nb", "wolfville, ns"))
```

And, if you don't care to modify the values yourself to zoom in or out, you can use `zoombbox()`.

```{r}
bb <- makebbox(50.1232, -122.9574, 50.1035, -123.0042)
zoombbox(bb, 0.8) # zooms out a tad
```

## Geocoding tools

Sometimes it's nice to be able to put cities on maps without having to look up their latitude and longitude, which is where the `geocode()` function comes in handy. It should be noted that the [ggmap package](https://cran.r-project.org/package=ggmap) also has a `geocode()` function that is also quite good, although didn't serve my purposes of calculating bounding boxes. If you do a lot of this, please [get your own API key](https://app.pickpoint.io/sign-up) so not everybody is using up mine (Google doesn't require one, although if you have an enterprise account you can pass a `key` just as you would for pickpoint).

```{r, message=FALSE}
results <- geocode(c("halifax, NS", "wolfville, NS", "moncton, NB"))
names(results) # several standard columns
results[c("query", "address", "lon", "lat")]
```

## Mapping tools

There's also a few mapping tools in prettymapr, most notably the `prettymap()` function, which evaluates its first argument while setting the margins to 0 (which is probably what you want for plotting a map). It also automatically adds a scalebar (although you could do this yourslef using the `addscalebar()` function).

```{r, message=FALSE, warning=FALSE}
library(rosm)
prettymap(osm.plot(searchbbox("nova scotia")))
```

Any plotting code has to go within `prettymap()`, so if you need multiple lines you can enclose them in curly brackets (`{}`). There's tons of options for `prettymap()` that let you customize the north arrow, scale bar etc., which you can find in the [prettymapr manual](https://cran.r-project.org/web/packages/prettymapr/prettymapr.pdf).

```{r, message=FALSE, warning=FALSE}
cities <- geocode(c("halifax, NS", "wolfville, NS", "moncton, NB"))
prettymap({
  osm.plot(searchbbox("nova scotia"))
  osm.points(cities$lon, cities$lat, pch=18)
})
```

Any plotting code that uses base plotting can be used inside `prettymap()`, including [rosm](https://cran.r-project.org/package=rosm), [rcanvec](https://cran.r-project.org/package=rcanvec), [OpenStreetMap](https://cran.r-project.org/package=OpenStreetMap), [cartography](https://cran.r-project.org/package=cartography), [marmap](https://cran.r-project.org/package=marmap), and others. If only a scale bar or north arrow are required, there are independent functions to add to existing code.

```{r, message=FALSE, warning=FALSE}
library(maptools)
data("wrld_simpl")
plot(wrld_simpl, xlim=c(-131, -52), ylim=c(10, 55))
addscalebar()
addnortharrow()
```

That's it for now! More convenience functions on the way...
